|

### 2021-03-26 paper/v2_taylor_francis/response_reviewers.md

---
output:
  pdf_document: default
bibliography: bibliography.bib
---

Dear Prof. Bian,

Thank you for providing the opportunity to revise and resubmit our paper.
We found the reviewer comments very useful and constructive and we believe they 
enabled us to improve the paper and clarify its findings.
We were also very pleased to see that the reviewers found our paper a timely and
important topic of research, and that we have addressed all their questions and 
concerns below and in the paper itself.

@@

We have added the following sentence to the first paragraph in section 3.2, explaining
why we believe any bias balances out:
"Indeed, those who seek to test their broadband are most likely to do so because
they are experiencing slower speeds than expected, although any skew towards slower
speeds is balanced by the likelihood that those who test their broadband are also
more 'tech-savvy' and / or have purchased higher speed packages that are not
delivering the promised level of service."

As we explain, our intentions were to analyse a dataset of experienced internet speeds.
Those who use the speedcheck service are not solicited in any way, and no personally
identifying data is included other than the geolocation of the test.

> 2. The results for network speed for the different regions are not very different with 1-2 out of 10Mbps difference and probably can not reflect the social disparities. Not sure if this is the facts in U.K. If so, then it's not a problem?

Firstly, mean speeds are not likely to be that different between regions of the UK
(e.g. North or South England). The greater differences are between urban, built-up
areas, and more rural areas within regions, although there are some local initiatives
to address this in some areas. However, we assume that your comment refers to Table 1
and the differences in mean network speed between clusters? Whilst these differences
are not large, the range for upload speeds is smaller than for download speeds,
and a difference of 1-2Mbps slower may well be noticeable at the wrong moment.
Importantly, we only consider the tests which took place during working hours, so
when speed drops, let's say, during a work Zoom call or when uploading large files,
it will be noticeable. This is why we created the clusters by how the time series
of speeds changes by day of the week and hour of the day, rather than focusing our
analysis on mean speeds for the period as a whole.

@@

"Such digital intersections and efficiencies have also changed 
transportation, connectivity and the structure of cities during the pandemic.
We now have enough data to confirm the drastic alteration observed in the space-time 
geography of cities around the world in the last year [e.g. @google2020; @shibayama2021impact]. 
Yet, although there is a broad agreement that these changes 
during the pandemic have played a pivotal role in stopping the spread of the virus 
[@jia2020population; @yang2020taking; @mu2020interplay], the extent to which the 
increased levels of working from home, the consequent decrease in commuting flows, 
and the altered structure of cities will remain post-pandemic is the subject
of considerable debate in the literature."

> 6. A review of cyberinfrastructure and how network varies in different region would be useful.

In section 2.2 of the literature review, we have added additional detail specific to the UK:

"This finding broadly applies in the UK, where studies that also analysed broadband
speed checks concluded that average speeds are lower in rural areas, something that
has not been improved by policy measures to increase competition 
[@riddlesden2014broadband; @nardotto2015unbundling]. In contrast,
@riddlesden2014broadband found that levels of deprivation did not correlate with
first level digital divides".

@@

> 2. Another major comment is on the multinomial logit regression (MLR). This is partly related to my first comment on a larger k. While there is no hard limit on the number of categories in MLR, the interpretation of results will be difficult with a high k as well. In addition, if the dependence among independent variables (collinearity) occurs, the estimation of the multinomial logit model parameters becomes inaccurate. According to Table 2, it is very likely some of the variables could present collinearity (e.g., distance to nearest metro and tech jobs).

Following your suggestion, we have now opted to present the cluster analysis and
the subsequent multinomial regression using $k = 9$. We indeed believe that the
results are more clear now. we have added the following justification in the Methods
and Data section:

"We initially run the algorithm for $k \in \mathbb{N} \bigcap [5,15]$, calculated
the cluster validity indices (CVIs) and then run the
subsequent multinomial regression -- see the end of this section for more details
for the different *k*.
Following @sardatime, to identify the optimal *k* we used the majority vote for
the following CVIs: Silhouette (max), Score function (max), Calinski-Harabasz (max),
Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).
Nevertheless, we opted against using the optimal $k = 13$ solution as it was too
large to allow for communicable LAD clusters. Instead we opted for a smaller $k = 9$,
which led to a rather similar spatial pattern and, importantly, to a much higher
R-squared in the subsequent explanatory regression ($0.44$ instead of $0.34$)."

Regarding multicolinearity, the highest correlations were observed between $NVQ4+$, 
which is a measure of being educated to degree level or higher,
and $professional jobs$ ($0.78$) and $NVQ4+$ and earnings ($0.69$). Hence, we decided
to exclude $NVQ4+$ from the regression.

@@

> 3. Although Fig. 1 and 2 exhibit different speed tests patterns in 2019 and 2020, the data mining outcome lacks a baseline (prior to 2020) to compare with. It is very likely that the clusters in 2019 would be different from that in 2020. It will make the results more robust. Of course, I will not push this given the page limit. But some insights may be helpful in the Discussion (however, a Discussion section is missing).

This is useful. We have added a new Appendix 2, which plots the clusters for 
the same period in 2019. The picture is very different, which is an indication of 
the changes in the temporal profiles of internet usage during the pandemic, as 
we say towards the end of section 4.1:

"It is worth noting here that when we apply the same methods to upload speed data for 
the same time period in 2019, the spatial distribution and composition of the nine 
clusters of LADs are very different -- see Appendix [2](#appendix2). This 
difference is an indication of the changes in the temporal profile of internet usage 
that took place during the pandemic, and which would be independent of any wider
trend in improving internet services."

@@

> 4. No discussion was included in this manuscript. I expect some insights into various important issues, such as problem causes, any broadband policies and socioeconomic solutions, equitable development, and research uncertainties.

We have tightened up the results section, which had to be re-written for the nine
clusters instead of 13, and begin our discussion of insights in the last paragraph 
of the results section. We then expand the conclusion section to include
a more in-depth discussion of how the results link to some of the issues you 
suggest.

We mention some uncertainties or avenues for future research in this final section,
but also see our response to Reviewer 1, point 1 for an explanation of biases.

@@

> 5.  There are three levels of digital divide. According to the last paragraph in Section 2, only the intersection of level-1 and level-3 was included in this research. I am wondering how level-2 may fit in the general research framework. Could education and demographic variables be used to represent level-2 digital divide?

We clarify that the occupations included in the regression are more about the level-2
digital divide, providing an indication of not only education, but digital
skills. The average earnings and, particularly during the pandemic, the new variable 
-- % furloughed, are indications of level-3 digital divides. We did have education
in the previous submission (NVQ4+), but have removed it due to multicollinearity
-- see response for point 2.

### 2021-03-26 paper/v2_taylor_francis/broadband.speed.covid.Rmd

Of the smaller clusters, cluster $7$ experiences quite similar mean speeds to 
cluster $9$, and a not too dissimilar standard deviation, but mean speeds in the 
morning peak are almost $10$% lower than the evening peak.
Cluster $3$ boasts speeds almost $1$Mb/s higher again, but suffers from even greater
slowdown in the morning peak. In comparison, clusters $2$ and $5$, home to a 
little less than half a million people, experience not only above average mean 
speeds, but also faster speeds in the morning peak compared to the evening.
Thus, in the five LADs in these two clusters, the temporal profile of internet use 
may be closer to what might have been expected pre-pandemic.
In comparison, the $15$ LADs in clusters $1$ and $8$, home to over $1.4$ million 
people, not only experience fairly average upload speeds and high ratios of 
standard deviation to the mean, but also experience much lower speeds during 
the morning peak than the evening peak. From the large spikes and dips shown on
Figure \ref{UpCluster}, it is likely that this poor reliability or consistency of 
morning internet speeds was more noticeable to those in cluster $1$, whilst those 
in cluster $8$ experienced the lack of reliable service as a problem throughout 
the day.

It is worth noting here that clusters of LAD based on the upload speeds are very 
different to the ones occur when we apply the same methods to upload speed data
for the same time period in 2019 -- see Appendix [2](#appendix2). This drastic 
difference between the LAD upload speed profiles before and during the pandemic is
an indication of the changes in internet usage that took place during the pandemic.

>>>>

Of the smaller clusters, the $15$ LADs in clusters $1$ and $8$, home to over $1.4$ 
million people, not only experience fairly average upload speeds and high ratios of 
standard deviation to the mean, but also experience much lower speeds during 
the morning peak than the evening peak. From the large spikes and dips shown on
Figure \ref{UpCluster}, it is likely that this poor reliability or consistency of 
morning internet speeds was more noticeable to those in cluster $1$, whilst those 
in cluster $8$ experienced the lack of reliable service as a problem throughout 
the day. Meanwhile, cluster $7$ experiences quite similar mean speeds to 
cluster $9$, and a not too dissimilar standard deviation, but mean speeds in the 
morning peak are almost $10$% lower than the evening peak.
Cluster $3$ boasts speeds almost $1$Mb/s higher again, but suffers from even greater
slowdown in the morning peak. In comparison, clusters $2$ and $5$, home to a 
little less than half a million people, experience not only above average mean 
speeds, but also faster speeds in the morning peak compared to the evening.
Thus, in the five LADs in these two clusters, the temporal profile of internet use 
may be closer to what might have been expected pre-pandemic.

It is worth noting here that when we apply the same methods to upload speed data for 
the same time period in 2019, the spatial distribution and composition of the nine 
clusters of LADs are very different -- see Appendix [2](#appendix2). This 
difference is an indication of the changes in the temporal profile of internet usage 
that took place during the pandemic, and which would be independent of any wider
trend in improving internet services.
